Velocity Map Imaging Spectroscopy of C$_2$H$^-$ and C$_2$D$^-$: a benchmark study of vibronic coupling interactions

### 2022-06-24 Resubmission/Laws_C2H_v2.tex
update resubmission

\begin{abstract}
High-resolution velocity-map imaged photoelectron spectra of the ethynyl anions C$_2$H$^-$ and C$_2$D$^-$ are measured over a range of photodetachment wavelengths, to investigate the complex interactions between the close lying $\tilde{X}^2\Sigma^+$ and $\tilde{A}^2\Pi$ electronic states. Vibronic coupling calculations are performed by transforming to a quasidiabatic representation, parametrised by \emph{ab-initio} CFOUR calculations, to model the interplay between pseudo Jahn-Teller and Renner-Teller coupling. This approach is combined with electron anisotropy measurements to assign all of the observed vibronic structure, providing a framework that may be used to accurately simulate spectra of larger C$_{2n}$H monohydride carbon chains.
\end{abstract}

@@

\section{Introduction}
Carbon monohydrides C$_{2n}$H, are a class of linear radicals that play an important role in combustion and interstellar chemistry~\cite{kie90,kie92,bou96,wil91,vui01}. These carbon chains have been observed in many interstellar environments, including planetary atmospheres~\cite{vui01,wil03,dob16}, comets~\cite{jac96}, dark clouds~\cite{ziu82,gup09}, and during heavy star formation~\cite{beu08,jia15}. Due to their relatively high abundance in a variety of astronomical conditions, they are believed to be promising candidates for Diffuse Interstellar Band (DIB) carriers~\cite{dou77,ful93,wat94,ful00,sch05,gup09}. The corresponding anions C$_{2n}$H$^-$ are also believed to play an important role in interstellar chemistry~\cite{mil17,gup09}. C$_6$H$^-$ was the first negative ion detected in space, after it was observed in two distinct interstellar molecular clouds IRC+10216 and TMC-1 in 2006~\cite{mcc06}. More recently, anions C$_4$H$^-$~\cite{cer07} and C$_8$H$^-$~\cite{bru07,rem07} have also been detected in a range of environments, include dark clouds~\cite{cor13}, prestellar cores,~\cite{sak10} and prostellar envelopes~\cite{sak07}.

Chain growth of the C$_{2n}$H radicals occurs predominately through acetylene addition,~\cite{wil03}
\begin{equation}
\text{C}_{2n}\text{H} + \text{C}_2\text{H}_2 \rightarrow \text{C}_{2n+2}\text{H}+\text{H}_2.
\end{equation}
Conversely, the formation of anions in space is believed to be driven by radiative electron attachment, charge transfer, and dissociative electron attachment. Current astrochemical modelling of interstellar clouds, such as IRC+10216, typically underestimate the observed abundances of CH$_4^-$ ~\cite{mil17,cor13,her08}. Therefore, it has been proposed that ion-neutral reactions
\begin{equation}
\text{C}_{2n}\text{H}^- + \text{C}_2\text{H}_2 \rightarrow \text{C}_{2n+2}\text{H}^- + \text{H}_2.
\end{equation}
may be significant chain growth mechanisms that also needs to be included in astrochemical models~\cite{mil17,bas19}. Studies of the extraterrestrial planetary atmosphere of Saturn's moon Titan suggest that both C$_2$H and C$_2$H$^-$ play an important role in the atmospheric chemistry~\cite{dob16,vui09,des17,vri18}. Multiple reaction pathways involving the C$_2$H$^-$ have been proposed, including reactions with C$_2$H$_2$ and HCN.

Our knowledge of interstellar chemistry relies critically on theory, to provide a link between astronomical observations and terrestrial laboratory studies. Microwave spectroscopy has successfully been used in this fashion to identify a large number of molecules in the interstellar medium (ISM). However for some species, particularly those with low abundances or no dipole moment, UV/vis spectroscopic methods are required for identification~\cite{mai97}. This creates a challenge for theory, as electronic spectral calculations are more difficult than those employed for pure rotational spectra~\cite{for10}. This becomes even more challenging when vibronic coupling effects are introduced, as they can be exceedingly difficult to model.

@@

In this work the vibronic spectrum of the ethynyl radical C$_2$H is examined with High-Resolution Photoelectron Imaging (HR-PEI) together with CFOUR~\cite{dev20}  \emph{ab-initio} vibronic-coupling calculations. C$_2$H is reported to be one of the most abundant molecules in the universe, and is the most thoroughly studied of the C$_{2n}$H species~\cite{wil91,hei99,for10}. Many different experimental techniques have been employed to help understand the complexities of the vibronic spectrum, including electron spin resonance~\cite{coc64,jin85}, laser magnetic resonance~\cite{bro88,pfe96,sch98}, microwave and millimeter wave spectroscopy~\cite{sas81,got83,end89,mul00}, infrared (matrix isolation~\cite{she87,jac87,for95} and Fourier Transform emission~\cite{ver88}) spectroscopy, laser induced fluorescence spectroscopy~\cite{hsu93,hsu95,chi99} and photoelectron spectroscopy of the negative ion~\cite{erv91,tay98,zho07}. C$_2$H has also received extensive theoretical attention to interpret the experimental results, however the large number of possible vibronic levels means this has proven to be a challenge~\cite{per90,per91,per91b,per91c,per92,car00,tar03,tar04,for10,sta21,gul21}. In this work we demonstrate how the construction of a quasidiabatic Hamiltonian allows the strength of vibronic interactions between coupled surfaces near a conical intersection to be estimated, sufficient to accurately simulate electronic and vibronic spectra. Anion HR-PEI spectroscopy maps both of the $^2\Sigma^+$ and $^2\Pi$ surfaces on an equal footing from the anion $\tilde{X}^1\Sigma^+$ state, allowing a direct comparison with \emph{ab-initio} modelling.

\section{Methods}
\subsection{Experimental Procedure}
Details of the HR-PEI spectrometer are given in Refs~\onlinecite{cav07} and~\onlinecite{dev17}. Briefly, C$_2$H$^-$ and C$_2$D$^-$ anions are produced by passing pure ethylene (C$_2$H$_4$ or C$_2$D$_4$) gas at $\sim2$ bar pressure through a pulsed solenoid valve, into the source vacuum chamber. The gas undergoes adiabatic supersonic expansion into a high-voltage discharge which generates negative ions. The anions are extracted, accelerated to 500~eV, and focussed into a novel gating, bunching, and re-referencing unit~\cite{ded01}, before being mass separated over a 2m time-of-flight region. The anion of interest (C$_2$H$^-$ or C$_2$D$^-$) is then isolated by an electrostatic gate. The ion packet is crossed with a tuneable detachment laser beam, generated from a Continuum Sunlite EX optical parametric oscillator pumped by the third harmonic of a Continuum Powerlite 9010 Nd:YAG laser. The laser output is then doubled by passing through a BBO crystal, to produce UV laser pulse energies between 1-5mJ at 10Hz. The wavelength of the laser light is measured using a HighFinesse WS7 UV wavemeter.

A modified velocity-map imaging lens images the detached photoelectrons to a 75mm diameter MCP/phosphor screen detector. Events are imaged by a 2048x2048 monochrome CCD camera (PCO 2000), with each frame transferred to a computer at a 10Hz repetition rate, and processed in real time to identify electron events, centroiding each position to
a subpixel accuracy. The electron positions are then written to a data file for subsequent analysis. The velocity-map image is centred and then circularized by an angular dependent-radial scaling determined by comparing adjacent radial slice intensity profiles~\cite{gas17}. An inverse Abel transformation of the VMI, based on the algorithm of Hansen and Law~\cite{han85,hic19}, returns a slice image of the 3D electron source distribution. Absolute energy calibration of the photoelectron spectra is achieved using published measurements of species, including C$_2^-$~\cite{law19b} and NO$_2^-$~\cite{law19}, that have been studied under similar conditions and photon energies as those used for the C$_2$H$^-$ measurements.

\subsection{Computational Details}
Quantum chemical computations were preformed using the CFOUR computational package~\cite{dev20}. The equation-of-motion ionised-potential coupled-cluster singles, doubles, and triples (EOM-IP-CCSDT) level of theory with an atomic natural orbital basis set of quadruple-zeta quality (ANO2) was used on a closed-shell anionic reference to perform geometry optimisations and frequency calculations (including 1h + 2h1p + 3h2p excitations~\cite{mat16}). EOM is a highly accurate and robust framework that provides access to different types of target electronic states, including ionised states~\cite{gul21b,and22,bar07}. The linear diabatic coupling constants were calculated analytically at the EOM-IP-CCSD/ANO2 level.

\section{Results and Analysis}
Ethynyl ions were produced in a pulsed-jet discharge of pure ethylene (C$_2$H$_4$) gas, and subsequently mass isolated via time of flight. Electrons were detached using a tuneable Continuum Sunlite Ex Optical Parametric Oscillator (OPO) pumped with the third harmonic (355~nm) of a Nd:YAG laser. Higher electron counts were also obtained with the direct use of the third and fourth harmonics (266~nm) of the Nd:YAG laser. Detached electrons are mapped onto a micro channel plate detector using a Velocity Map Imaging (VMI) lens. An illustrative velocity map image containing $\sim$4 million electrons, collected from 355~nm (3.49~eV) photodetachment of C$_2$H$^-$ is shown in Fig.~\ref{fig:1}(a). The electrons are distributed radially according to their speed, with slow electrons near the image center, and fast electrons located towards the outer edge. Due to the relatively large C$_2$H electron affinity (EA) of 2.969~eV~\cite{erv91}, photodetachment at 355~nm produces photoelectrons with small (< 0.5~eV) kinetic energies. These slow electrons allow the use of a low repeller voltage ($-600$~V) for the VMI lens, spreading the electron velocity distribution across the whole detector to give a high electron velocity resolution.

Despite the photodetachment energy being close to threshold, two electronic states of neutral C$_2$H are observed. The faster electrons, on the outer edge, correspond to C$_2$H$(\tilde{X}\,^2\Sigma^+)+e^- \leftarrow $C$_2$H$^-(\tilde{X}\,^1\Sigma^+)+h\nu$ photodetachment, and these are preferentially distributed around the poles of the image, indicative of a positive anisotropy parameter. Conversely the angular distribution of the slower electrons near the center is skewed towards the equator, indicative of a negative anisotropy parameter~\cite{buc70}. These electrons may be assigned to photodetachment to the first excited state C$_2$H$(\tilde{A}\,^2\Pi)+e^- \leftarrow $C$_2$H$^-(\tilde{X}\,^1\Sigma^+)+h\nu$.

\begin{figure}
	\includegraphics[width=0.8\textwidth]{figures/Fig1}
	\caption{(a) Velocity map image of electrons from C$_2$H$^-$ photodetachment at 355~nm. (b) Velocity map image of C$_2$D$^-$ at 355~nm. Both images contain $\sim4$ million electrons. (c) Comparison of the corresponding 355~nm photoelectron spectra for C$_2$H$^-$ and C$_2$D$^-$. Structure below $\sim 27,000~$cm$^{-1}$ is associated with the ground state $\tilde{X} ^2\Sigma^+$, while structure above  $\sim 27,000~$cm$^{-1}$ is largely due to the excited neutral $\tilde{A} ^2\Pi$ state.}
	\label{fig:1}
\end{figure}

To investigate the vibronic coupling interaction between these two nearby $\Sigma^+$ and $\Pi$ surfaces, deuterated ethynyl C$_2$D$^-$ was also studied. The C$_2$D$^-$ ions were produced in a discharge of pure C$_2$D$_4$ gas and measurements were performed under the same experimental conditions as those for C$_2$H$^-$. An illustrative velocity-map image of photodetachment at 355~nm from C$_2$D$^-$ is shown in Fig.~\ref{fig:1}(b). Similarly to Fig~\ref{fig:1}(a) two electronic transitions are observed. While the distribution of fast electrons is similar in both images, a striking difference is observed for the slow electrons, near the detector center. In Fig.~\ref{fig:1}(a) a series of weak rings are observed, however in Fig.~\ref{fig:1}(b) the image is dominated by two distinct rings.

The velocity-map images of Fig.~\ref{fig:1} were inverted using the Abel inversion methods detailed in PyAbel~\cite{hic19} to extract the corresponding photoelectron spectra, presented in Fig.~\ref{fig:1}(c). Below $27,000~$cm$^{-1}$ binding energy, corresponding to the $\tilde{X} ^2\Sigma^+$ surface ,the C$_2$H$^-$ and C$_2$D$^-$ spectra are similar in structure. Both are dominated by the origin transition, shifted by $\sim10~$cm$^{-1}$ in the deuterated spectrum, with progressions involving the v$_2$($\pi$) bending and v$_3$($\sigma$) CC stretch vibrational normal modes. This structure also includes transitions involving an odd quanta of v$_2$ bending excitation ($2^{n+1}$) which are totally forbidden within the Franck-Condon approximation. The presence of these transitions in the spectra is an indicator of Herzberg-Teller (HT) vibronic coupling between the ground $^2\Sigma^+$ and nearby excited $^2\Pi$ electronic surfaces, as
\begin{equation}
\Sigma^+ \otimes \pi = \Pi. 
\end{equation}

Above $27,000~$cm$^{-1}$, near the $\tilde{A}^2\Pi$ state origin, significant differences are observed between the C$_2$H$^-$ and C$_2$D$^-$ photoelectron spectra. In C$_2$H$^-$, 5 sharp peaks are observed, spaced by $\sim95~$cm$^{-1}$. However in the deuterated spectrum, one dominant peak is observed at $27,792$~cm$^{-1}$, with 3 weaker peaks centred around $27,360~$cm$^{-1}$. Unlike the structure below $27,000$~cm$^{-1}$ none of these peaks are able to be readily assigned to vibronic transitions, due to the presence of strong coupling interactions between the nearby $\Sigma^+$ and $\Pi$ surfaces.

\subsection{Vibronic Coupling Interactions}
The complex spectral structure observed near the $\tilde{A}^2\Pi$ state origin in Fig.~\ref{fig:1}(c) may be understood by considering the $v_2$ bending vibrational mode. To account for the degeneracy of this mode, the vibronic quantum number $\ell_i$ may be introduced, representing the angular momentum associated with the bending motion. This may take a value of $\ell_i = v_i, v_i-2, v_i-4,\dots,1$ or $0$, where $v_i$ is the quanta of bending excitation. In the Born-Oppenheimer approximation, different vibronic energy levels $\ell_i$ are degenerate, however in cases with strong rovibronic coupling, this degeneracy in $\ell_i$ is lost.

@@

These interactions spread the electronic orgin of the $\tilde{A}\,^2\Pi$ state over several vibronic levels. Therefore, instead of assigning a defined origin, the observed peaks in the spectrum may be assigned to coupled admixtures of vibronic transitions involving the three potential energy surfaces,
\begin{equation}
\Psi_f = \sum_\xi \psi_e^\xi \sum_k C_{fk}^\xi\phi_{fkm}^\xi,
\label{eq:teller3} 
\end{equation}
where $\psi_e^\xi$ is the diabatic electronic wavefunction, and $\phi_{fkm}^\xi$ is the spin-rovibrational wavefunction. $\xi$ represents the electronic states used in the expansion ($\xi=\Sigma^+(1A'),~\Pi^+(2A'),~\Pi^-(1A'')$). A depiction of these three interacting surfaces, generated from the parameters in Ref.~\onlinecite{tar03}, is given in Fig.~\ref{fig:2}.

\begin{figure}
	\centering
	\includegraphics[width=0.5\textwidth]{figures/Fig2}
	\caption{Adiabatic potential energy surfaces of $\tilde{X}^2\Sigma^+$, $\tilde{A}^2\Pi^-$, and $\tilde{A}^2\Pi^+$ states, calculated using the parameters calculated by Tarroni and Carter~\cite{tar03}.}
	\label{fig:2}
\end{figure}

The adiabatic potential energy surfaces in Figure~\ref{fig:2} were evaluated from the variational parameters of Tarroni and Carter~\cite{tar03}, illustrating the loss of degeneracy for the doublet $^2\Pi$ surfaces, due to the Renner-Teller and pseudo Jahn-Teller interactions. This technique has been used to calculate a multitude of admixed vibronic levels. More than 100 possible C$_2$H vibronic states have been calculated, up to 10,000~cm$^{-1}$ above the $\tilde{X} ^2\Sigma^+$ origin. However, assigning experimental spectra has remained a challenge, partly due to the large number of calculated levels providing for multiple possible assignments for each experimentally observed transition.

\subsection{Coupling Calculations}
In order to guide spectral assignment of C$_2$H and future astronomical searches for other C$_n$H radicals, transition intensities are also needed. However, due to the strong vibronic coupling interactions, transition intensities can not be obtained via simple Franck-Condon calculations based on quantum chemistry methods. To account for these interactions, the photoelectron spectrum of the C$_2$H$^-$ anion is simulated using a quasidiabatic Hamiltonian of the type introduced by Koeppel, Domcke and Cederbaum (KDC).~\cite{kou84,dom81} In this approach, the Hamiltonian is represented in a basis of quasidiabatic (slowly varying) electronic  states for which the kinetic energy operator can be assumed diagonal.~\cite{pac93} For C$_2$H, the KDC Hamiltonian comprises three states - the two components of the $^2\Pi$ state and the ground $^2\Sigma^+$ state - and is then projected onto a vibrational basis, usually chosen as a direct product of harmonic oscillators. Diagonalization of the corresponding matrix yields the molecular states (which are given in terms of a Born-Huang expansion), and the squared projections of the corresponding eigenvectors onto the ground state of the anion yield the relative intensities. The latter is true only if the photodetachment cross sections for the $^2\Sigma^+$ and $^2\Pi$ states are assumed equal, but different cross sections of the two states can be incorporated by scaling the intensities of states according to their (vibronic) symmetry.~\cite{sta11}

Details of the construction and parametrization of KDC Hamiltonians can be found elsewhere in the literature,~\cite{sta21,ich06,ich08,ich09,wei17} and the procedure followed will be discussed here only briefly. The present calculations use the so-called quadratic vibronic coupling (QVC) model.~\cite{ich06} For the system at hand, the QVC model Hamiltonian  assumes the following form

\begin{align}
H &= T_n1+V,\\
V &= \bordermatrix{ & X & A(a') & A(a'') \cr
	X &\substack{\Delta_0^X+F_1^Xq_1+F_3^Xq_3+\\ 
		\frac{1}{2} \sum_{ij}F^X_{ij}q_iq_j}& \lambda q_{2a} & \lambda q_{2b} \cr
	A(a') & \lambda q_{2a} & \substack{\Delta^A_0+F^A_1q_1+F^A_3q_3\\+\frac{1}{2}\sum_{ij}F^A_{ij}q_iq_j+\\\frac{1}{2}\eta(q_{2a}^2-q_{2b}^2)} & \eta q_{2a}q_{2b} \cr
	A(a'') & \lambda q_{2b} & \eta q_{2a}q_{2b} & \substack{\Delta_0^A+F_1^Aq_1+F_3^Aq_3\\+\frac{1}{2}\sum_{ij}F_{ij}^Aq_iq_j\\-\frac{1}{2}\eta(q_{2a}^2-q_{2b}^2) } } \qquad
\label{eq:calc1}
\end{align}

where the summations run over the dimensionless normal modes ($q_i$) that serve as the coordinate system for the problem.~\cite{sta21,car00} For convenience, the latter are chosen to be those of the anion, which considerably facilitates calculation of the spectral intensities.  In Eq.~(\ref{eq:calc1}), the diagonal terms of V (excluding those that carry the Renner-Teller coupling constant $\eta$) represent the quasidiabatic potential energy surfaces of the $^2\Sigma^+$ state, and the two components of the $^2\Pi$ state, chosen here as those in which the unpaired electron lies within or is perpendicular to an arbitrarily chosen plane, designated as A(a$'$) and A(a$''$), respectively. $\Delta_0^X$ is the separation of the anion and $^2\Sigma^+$ states at the origin of the coordinate system (the vertical electron detachment energy), and $\Delta_0^{A}$ is the gap between anion and $^2\Pi$ states at the same geometry.

For the modes of $\sigma$ symmetry ($q_1$ and $q_3$), the diabatic forces ($F_i$) and force constants ($F_{ij}$) coincide with those of the adiabatic potential energy surfaces.   However, for the bending mode, the diabatic (F$_{22}$) and adiabatic (f$_{22}$) force constants differ, and the parametrization is somewhat more involved.   For each component of the $^2\Pi$ state, either the 2a or 2b component of the bending vibration will maintain the A$'$ electronic symmetry that is needed to couple with the $^2\Sigma^+$ state.  Designating these as 2a for the A(a$'$) state and 2b for the A(a$''$) state (as is implicit in Eq.~\ref{eq:calc1}), the diabatic force constants for the bending mode in the $^2\Sigma^+$ and $^2\Pi$ states can be written as

\begin{align}
F_{22}^X &= f_{22}^X+\frac{2\lambda^2}{(\Delta_0^A-\Delta_0^X)}\\
F_{22}^A &= f_{2a2a}^{A(a')}-\frac{2\lambda^2}{(\Delta_0^A-\Delta_0^X)}-\eta
\label{eq:calc2}
\end{align}

where the Renner-Teller interaction strength ($\eta$) is determined from

\begin{equation}
\eta = \frac{1}{2}\left[ f_{2a2a}^{A(a')}-f_{2b2b}^{A(a')}-\frac{2\lambda^2}{(\Delta_0^A-\Delta_0^X)}\right]
\label{eq:calc3}
\end{equation}

once the interstate coupling ($\lambda$ which is calculated analytically in this work, see below) is known.

When the potential in Eq.~(\ref{eq:calc1}) is diagonalized, the adiabatic states that are used for its parametrization are precisely recovered through terms second order in displacement. The parametrization of the potential in this work was completed using the CFOUR computational package~\cite{dev20} and the quasidiabatic ansatz of Ichino~\emph{et al.}~\cite{ich09} Briefly,the EOM procedure is used to operationally define quasidiabatic states (those that relax according to a well-behaved reference state wavefunction, which is that of the anion level) and the coupling constants are then evaluated as the first derivative of the off-diagonal elements of the electronic Hamiltonian in the basis defined by this representation. Vertical excitation energies ($\Delta_0$), and linear diabatic force and coupling constants between the electronic states of interest were calculated analytically using the highly accurate EOM-IP-CCSDT/ANO2 procedure. The interstate coupling constants ($\lambda$) were calculated analytically at the EOM-IP-CCSD/ANO2 level. All calculated values used to parametrize the KDC potential in Eq.~(\ref{eq:calc1}) are presented in Table~\ref{tab:1}.

>>>>

\begin{abstract}
High-resolution velocity-map imaged photoelectron spectra of the ethynyl anions C$_2$H$^-$ and C$_2$D$^-$ are measured at photon wavelengths between $355-266$~nm, to investigate the complex interactions between the close lying $\tilde{X} ^2\Sigma^+$ and $\tilde{A} ^2\Pi$ electronic states. An indicative kinetic energy resolution of 0.4\%, together with full angular dependence of the fast electrons, provides a detailed description of the vibronically coupled structure. It is demonstrated that a modest quadratic vibronic coupling model, parametrized by the quasidiabatic ansatz, is sufficient to accurately recreate all of the observed vibronic interactions. Simulated spectra are shown to be in excellent agreement with the experimental data, verifying the proposed model, providing a framework that may be used to accurately simulate spectra of larger C$_{2n}$H monohydride carbon chains. New spectral assignments are supported by experimental electron anisotropy measurements and Dyson orbital calculations.
\end{abstract}

@@

\section{Introduction}
Carbon monohydrides C$_{2n}$H, are a class of linear radicals that play an important role in combustion and interstellar chemistry~\cite{kie90,kie92,bou96,wil91,vui01}. These carbon chains have been observed in many interstellar environments, including planetary atmospheres~\cite{vui01,wil03,dob16}, comets~\cite{jac96}, dark clouds~\cite{ziu82,gup09}, and during heavy star formation~\cite{beu08,jia15}. Due to their relatively high abundance in a variety of astronomical conditions, they are believed to be promising candidates for Diffuse Interstellar Band (DIB) carriers~\cite{dou77,ful93,wat94,ful00,sch05,gup09}. The corresponding anions C$_{2n}$H$^-$ are also believed to play an important role in interstellar chemistry~\cite{mil17,gup09}. C$_6$H$^-$ was the first negative ion detected in space, after it was observed in two distinct interstellar molecular clouds IRC+10216 and TMC-1 in 2006~\cite{mcc06}. More recently, anions C$_4$H$^-$~\cite{cer07} and C$_8$H$^-$~\cite{bru07,rem07} have also been detected in a range of environments, including dark clouds~\cite{cor13}, prestellar cores,~\cite{sak10} and prostellar envelopes~\cite{sak07}.

Chain growth of the C$_{2n}$H radicals occurs predominately through acetylene addition,~\cite{wil03}
\begin{equation}
\text{C}_{2n}\text{H} + \text{C}_2\text{H}_2 \rightarrow \text{C}_{2n+2}\text{H}+\text{H}_2.
\label{eq:1}
\end{equation}
Conversely, the formation of anions in space is believed to be driven by radiative electron attachment, charge transfer, and dissociative electron attachment. Current astrochemical modelling of interstellar clouds, such as IRC+10216, typically underestimate the observed abundances of CH$_4^-$ ~\cite{mil17,cor13,her08}. Therefore, it has been proposed that ion-neutral reactions
\begin{equation}
\text{C}_{2n}\text{H}^- + \text{C}_2\text{H}_2 \rightarrow \text{C}_{2n+2}\text{H}^- + \text{H}_2
\label{eq:2}
\end{equation}
may be significant chain growth mechanisms that should be included in astrochemical models~\cite{mil17,bas19}. Studies of the extraterrestrial planetary atmosphere of Saturn's moon Titan suggest that both C$_2$H and C$_2$H$^-$ play an important role in the atmospheric chemistry~\cite{dob16,vui09,des17,vri18}. Many reaction pathways involving the C$_2$H$^-$ have been proposed, including reactions with C$_2$H$_2$ and HCN.

Our knowledge of interstellar chemistry relies critically on theory, to provide a link between astronomical observations and terrestrial laboratory studies. Microwave spectroscopy has successfully been used in this fashion to identify a large number of molecules in the interstellar medium (ISM). However for some species, particularly those with low abundances or no dipole moment, UV/vis spectroscopic methods are required for identification~\cite{mai97}. This creates a challenge for theory, as electronic spectral calculations are more difficult than those employed for pure rotational spectra~\cite{for10}. A challenge more arduous when vibronic coupling effects are present, as they can be exceedingly difficult to model.

@@

C$_2$H is reported to be one of the most abundant molecules in the ISM, and is the most thoroughly studied of the C$_{2n}$H species~\cite{wil91,hei99,for10}. Many different experimental techniques have been employed to help understand the complexities of the vibronic spectrum, including electron spin resonance~\cite{coc64,jin85}, laser magnetic resonance~\cite{bro88,pfe96,sch98}, microwave and millimeter wave spectroscopy~\cite{sas81,got83,end89,mul00}, infrared (matrix isolation~\cite{she87,jac87,for95} and Fourier Transform emission~\cite{ver88}) spectroscopy, laser induced fluorescence spectroscopy~\cite{hsu93,hsu95,chi99} and photoelectron spectroscopy of the negative ion~\cite{erv91,tay98,zho07}. Zhou~\emph{et al.} measured photodetachment of C$_2$H$^-$ and C$_2$D$^-$ to both the $\tilde{X} ^2\Sigma^+$ and $\tilde{A} ^2\Pi$ surfaces using slow electron velocity-map imaging (SEVI), which revealed detailed structure around the $\tilde{A} ^2\Pi$ state origin~\cite{zho07}. Comparison with calculations by Tarroni and Carter~\cite{tar03} resulted in detailed spectral assignments, however this was limited by the suppression of $p-$wave structure near threshold.

\subsection{Vibronic Coupling Interactions}
The complex spectral structure observed near the $\tilde{A}^2\Pi$ state origin in previous experiments may be understood by considering the $v_2$ bending vibrational mode. To account for the degeneracy of this mode, the vibronic quantum number $\ell_i$ may be introduced, representing the angular momentum associated with the bending motion. This may take a value of $\ell_i = v_i, v_i-2, v_i-4,\dots,1$ or $0$, where $v_i$ is the quantum of bending excitation. In the Born-Oppenheimer approximation, different vibronic energy levels $\ell_i$ are degenerate, however in cases with strong rovibronic coupling, this degeneracy in $\ell_i$ is lost.

@@

These interactions spread the electronic orgin of the $\tilde{A}\,^2\Pi$ state over several vibronic levels. Therefore, instead of assigning a defined origin, the peaks observed in experimental spectra may be assigned to coupled admixtures of vibronic transitions involving the three potential energy surfaces,
\begin{equation}
\Psi_f = \sum_\xi \psi_e^\xi \sum_k C_{fk}^\xi\phi_{fkm}^\xi,
\label{eq:3} 
\end{equation}
where $\psi_e^\xi$ is the diabatic electronic wavefunction, and $\phi_{fkm}^\xi$ is the spin-rovibrational wavefunction. $\xi$ represents the electronic states used in the expansion ($\xi=\Sigma^+(1A'),~\Pi^+(2A'),~\Pi^-(1A'')$).

C$_2$H has received extensive theoretical attention in an attempt to interpret the experimental results~\cite{per90,per91,per91b,per91c,per92,car00,tar03,tar04,for10,sta21,gul21}. This includes the variational method of Tarroni and Carter, which has been used to calculate a multitude of admixed vibronic levels in C$_2$H. A depiction of the three interacting potential energy surfaces in Eq.~(\ref{eq:3}), generated from the variational parameters of Ref~\onlinecite{tar03}, is shown in Fig.~\ref{fig:1}. This illustrates the loss of degeneracy for the doublet $^2\Pi$ surfaces along the CCH bending coordinate, due to the Renner-Teller and pseudo Jahn-Teller interactions. From these surfaces more than 100 possible C$_2$H vibronic states have been calculated, up to 10,000~cm$^{-1}$ above the $\tilde{X} ^2\Sigma^+$ origin~\cite{tar03}. However, assigning experimental spectra has remained a challenge, partly due to the large number of calculated levels, providing multiple options for each experimentally observed transition. The complexity of the variational approach also limits the extension of this method to larger C$_{2n}$H chains.

In this work we demonstrate that the construction of a quasidiabatic Hamiltonian allows the strength of vibronic interactions between coupled surfaces near a conical intersection to be estimated, sufficient to accurately simulate electronic and vibronic spectra. Anion HR-PEI spectroscopy maps both of the $^2\Sigma^+$ and $^2\Pi$ surfaces on an equal footing from the anion $\tilde{X}^1\Sigma^+$ state, allowing a direct comparison with the \emph{ab-initio} modelling. This is combined with experimental anisotropy measurements and Dyson orbital analysis to examine the vibronic symmetry of each observed transition to verify spectral assignments.

\begin{figure}
	\centering
	\includegraphics[width=0.5\textwidth]{figures/Fig1}
	\caption{Adiabatic potential energy surfaces of $\tilde{X}^2\Sigma^+$, $\tilde{A}^2\Pi^-$, and $\tilde{A}^2\Pi^+$ states, calculated using the parameters calculated by Tarroni and Carter~\cite{tar03}.}
	\label{fig:1}
\end{figure}

\section{Methods}
\subsection{Experimental Procedure}
Details of the HR-PEI spectrometer are given in Refs~\onlinecite{cav07} and~\onlinecite{dev17}. Briefly, ethynyl ions C$_2$H$^-$ and C$_2$D$^-$ were produced in a pulsed-jet discharge of pure ethylene (C$_2$H$_4$/C$_2$D$_4$) gas, and subsequently mass isolated over a 2m-long time of flight region~\cite{ded01}. Electrons were detached using a tuneable Continuum Sunlite Ex Optical Parametric Oscillator (OPO) pumped with the third harmonic (355~nm) of a Nd:YAG laser. Higher electron counts were also obtained with the direct use of the third and fourth harmonics (266~nm) from the Nd:YAG laser. The detached electrons are mapped onto a micro-channel plate detector using a modified Velocity Map Imaging (VMI) lens. Each detachment event is imaged by a 2048x2048 monochrome CCD camera (PCO 2000), with each frame transferred to a computer within the 10Hz repetition rate of the laser and processed in real time to identify electron events, centroiding each position to a subpixel accuracy.

The resulting velocity-map image is centred and then circularized by an angular dependent-radial scaling determined by comparing adjacent radial slice intensity profiles~\cite{gas17}. An inverse Abel transformation of the velocity-map image, based on the algorithm of Hansen and Law~\cite{han85,hic19}, returns a slice image of the 3D electron source distribution. Absolute energy calibration of the photoelectron spectra is achieved using published measurements of species, including C$_2^-$~\cite{law19b} and NO$_2^-$~\cite{law19}, that have been studied under similar conditions and photon energies as those used for the C$_2$H$^-$ measurements.

%--------------------------------------------------------------------

\subsection{Computational Details}
Quantum chemical computations were preformed using the CFOUR computational package~\cite{dev20}. The equation-of-motion ionised-potential coupled-cluster singles, doubles, and triples (EOM-IP-CCSDT) level of theory with an atomic natural orbital basis set of quadruple-zeta quality (ANO2) was used on a closed-shell anionic reference to perform geometry optimisations and frequency calculations (including 1h + 2h1p + 3h2p excitations~\cite{mat16}). EOM is a highly accurate and robust framework that provides access to different types of target electronic states, including ionised states, making it well suited to the study of photodetachment~\cite{gul21b,and22,bar07}. The linear diabatic coupling constants were calculated analytically at the EOM-IP-CCSD/ANO2 level.

Dyson orbitals, that are used to analyse the experimental electron anisotropy, were generated using Q-Chem software~\cite{sha15} at the EOM-CCSD/aug-gg-pVTZ level. Anisotropy parameters were then calculated using the ezDyson software of Gozem and Krylov~\cite{goz22}.

\subsection{The Quasidiabatic Ansatz} 
In order to guide spectral assignment of C$_2$H and future astronomical searches for other C$_n$H radicals, transition intensities are required. However, due to the strong vibronic coupling interactions, transition intensities can not be obtained via simple Franck-Condon calculations based on quantum chemistry methods. To account for these interactions, the photoelectron spectrum of the C$_2$H$^-$ anion is simulated, in this work, using a quasidiabatic Hamiltonian of the type introduced by Koeppel, Domcke and Cederbaum (KDC).~\cite{kou84,dom81} In this approach, the Hamiltonian is represented in a basis of quasidiabatic (slowly varying) electronic  states for which the kinetic energy operator can be assumed diagonal.~\cite{pac93} For C$_2$H, the KDC Hamiltonian comprises three states - the two components of the $^2\Pi$ state and the ground $^2\Sigma^+$ state - and is then projected onto a vibrational basis, usually chosen as a direct product of harmonic oscillators. Diagonalization of the corresponding matrix yields the molecular states (which are given in terms of a Born-Huang expansion), and the squared projections of the corresponding eigenvectors onto the ground state of the anion yield the relative intensities. The latter is true only if the photodetachment cross sections for the $^2\Sigma^+$ and $^2\Pi$ states are assumed equal, but different cross sections of the two states can be incorporated by scaling the intensities of states according to their (vibronic) symmetry.~\cite{sta11}
